---
title: "Does tipsy sommelier rate higher?"
author: "Damian Drewulski - damiannelus"
date: "12 maja 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Libraries initialization, echo=FALSE, warning=FALSE}
library("ggplot2")
library("plyr")
library("gridExtra")
library("scales")
library("GGally")
library("reshape")

```
```{r Reading data, echo=FALSE}
rm(list=ls())
wwhites <- read.csv2('C:/Users/Y50-70/Documents/wine_final_project/data/wineQualityWhites.csv', sep = ',')
wreds <- read.csv2('C:/Users/Y50-70/Documents/wine_final_project/data/wineQualityReds.csv', sep = ',')
wreds$.id <- "Red"
wwhites$.id <- "White"
w <- rbind(wreds,wwhites)
sel_cols = c(2:12)
as.numeric.factor <- function(x) {as.numeric(as.character(x))}
w[,sel_cols] <- lapply(w[,sel_cols], as.numeric.factor)
```

**Assumption**: to define alcohol categories, I used quantiles of binded data frames. Here are categories:
```{r Alcohol category extraction, echo=FALSE}
alcohol.breaks <- as.numeric(quantile(w$alcohol))
alcohol.breaks
w$alcohol.category <- cut(w$alcohol, breaks = alcohol.breaks, include.lowest = TRUE)
ggplot(w, aes(quality)) + geom_histogram(aes(fill = alcohol.category, y = (..count..)/sum(..count..)), binwidth = 0.5, position = "dodge") + scale_y_continuous(labels = percent) + facet_wrap(~.id, nrow=2)
```

However, the question from the first clue was a little puckish, it turns out, both for red and white wines, experts values more high-alcoholic wines.
At the same time, wines with low contamination of alcohol are popular among poor grades.

#### TODO: ADD numeric correlation.